---
title: "Slopes & Drainages"
output: html_notebook
---


```{r read in}
library(tidyverse)
library(ggplot2)
#setwd("<YOURPATHWAY>")
wd <- getwd()
new.data<- read.csv(paste(wd,'/Data/MetaData.csv',sep=""), header=TRUE, sep=",") #MetaData saved feb 11th
data<- read.csv(paste(wd,'/Data/Data.csv',sep=""), header=TRUE, sep=",")%>% #Data.csv from .xlsx feb 25th
  select(StudyID:Comments) 
r.squared<- read.csv(paste(wd,'/Data/TraitR2_among.csv',sep=""), header=TRUE, sep=",") #R2 generated feb 11th 
```

###Misc Notes on populations
North: 
Yarra, Marianne, Paria, Madamas, Damier, Brasso Seco, Balandra *Tompire*, *Ricon*, *Mission*, *Toco* (*northeast?*)

South: 
Caroni Drainage (Tunapuna, Lopinot, Guanapo, Aripo, Cuara, Tranquille, Caroni, Arima, Arouca, Mausica, Tacarigua, El Cedro, Ceniza, Naranjo, Maracas, Caroni Lower Dyke, Cuara, Manacal)
Oropuche Drainage (Quare, Oropuche, Turure, La Seiva, Rio Barro, Rio Grande, La Selva, Valencia, Matura)

*Questionable/Notes*
Madamas: Potentially a northern drainage, it's right beside the Paria to the east
Naranjo: upper aripo
Tumbason:south, drainage? not included anymore
Hollis: South? not included anymore
Tompire: west coast usually counted as northern range but not connected to others 
Mission: North but over near Tompire
Ricon: north but east and not connected to a drainage
Toco: north east like tompire
Cunupia Lower Dyke: near caroni but listed as distinct in 23, has gps in supplementary mat.	
Manzanilla:	No gps, listed as separate drainage
Matura: inlcuded for now in oropuche but some say it's not because its far east and north a bit
Cenzina changed to Ceniza (typo confirmed in paper)

*Southern trinidad outside northern mountain range*

Southern trinidad outside northern mountain range
Pilote and lower trib 1 &2 are a part of a Mayaro drainage, in southern trinidad)
Lizard
Guapo
Coffee
Visigney
Silverstream
Arena: Near the arena reservoir some place? thoroughly out of the mountains
Vance: South West (from jackie outside the northern range, also in paper 23, but not listed in supplementary material as a site??)

### Formatting columns
```{r population column}
data$temp<-NA #temp columns
data$pop<-data$PopulationRaw

names<-c("Arima","Arena","Arouca","Aripo", "Quare", "Yarra", "Tompire", "Marianne", "Paria", "El Cedro", "Madamas", "Oropuche", "Guanapo", "La Selva", "Turure", "Damier", "Tacarigua", "Mausica", "Manacal", "Ceniza", "Caroni", "Valencia", "Oropuna", "Tunapuna", "Lopinot", "Guapo", "Vance", "Visigney", "Pilote", "Lizard", "Silverstream", "Matura","Cunupia", "Coffee", "Maracas", "Caura", "Toco", "Manzanilla", "Mission", "Ricon", "Balandra", "Brasso Seco", "Rio Grande", "Rio Barro", "Caroni Lower Dyke", "Cunupia Lower Dyke")
for (i in names){ data<-mutate(data, pop = (ifelse(grepl(i, PopulationRaw), i,  temp)))
data$temp<-ifelse(data$pop==i, i, data$temp)} #replaces variations of populations with the root, ie Aripo 4 -> Aripo

data$pop<-ifelse(is.na(data$pop),data$PopulationRaw, data$pop) #fill NAs with the raw population data

#fix typos

data$pop<-gsub("ElCedro1", "El Cedro", data$pop)
data$pop<-gsub("ElCedro2", "El Cedro", data$pop)
data$pop<-gsub("Lower Guapo 1", "Guapo", data$pop)
data$pop<-gsub("Lower Guapo 2", "Guapo", data$pop)
data$pop<-gsub("Laseiva", "La Seiva", data$pop)
data$pop<-gsub("Cenzina", "Ceniza", data$pop)
data$pop<-gsub("Lower Trib 1", "Mayaro", data$pop)
data$pop<-gsub("Lower Trib 2", "Mayaro", data$pop)
data$pop<-gsub("Tompire 2", "Tompire", data$pop)
data$pop<-gsub("guanapo", "Guanapo", data$pop)
data$pop<-gsub("Lower La Seiva", "La Seiva", data$pop)
data$pop<-gsub("Upper turure", "Turure", data$pop)
data$pop<-gsub("Tacargua", "Tacarigua", data$pop)
data$pop<-gsub("Oropouche", "Oropuche", data$pop)
data$pop<-gsub("St Joseph", "Saint Joseph", data$pop)
data$pop<-gsub("Tac/Tig", "Tacarigua", data$pop) # Need to confirm this!!!!
#data$pop<-ifelse(grepl("Dyke", data$PopulationRaw), data$PopulationRaw, data$pop)

summary(as.factor(data$pop))
data$Population<-data$pop
data<-data %>% 
  select(StudyID:Comments)
```



```{r typos and river lists, eval=FALSE}
#fix typos not previously found
data$TraitType2<-gsub("Life-history", "Life history", data$TraitType2)
data$PopulationRaw<-gsub("Oropouche", "Oropuche", data$PopulationRaw)

data$Population<-as.factor(data$Population)
river.list<-data %>% 
  select(Population) %>% 
  unique()

North<-c("Yarra", "Marianne", "Paria", "Madamas", "Damier", "Brasso Seco", "Balandra") #I think we should keep these together for drainage level, logistically it'll be easier and its how others do it (paper 2 is an example)

Northeastern<- c("Tompire", "Ricon", "Mission", "Toco")

Other<-c("Pilote", "Lizard", "Guapo", "Coffee", "Silverstream", "Arena", "Vance", "Mayaro", "Cunupia Lower Dyke", "Manzanilla", "Visigney") #mostly southern range, jessica stephenson paper, but some others

South<-c("Tunapuna", "Lopinot", "Guanapo", "Aripo", "Caura", "Tranquille", "Caroni", "Arima", "Arouca", "Mausica", "Tacarigua", "El Cedro", "Ceniza", "Naranjo", "Maracas", "Caroni Lower Dyke", "Manacal", "Quare","Hollis", "Oropuche", "Turure", "La Seiva", "Rio Barro", "Rio Grande", "La Selva", "Valencia", "Matura", "Hollis", "Saint Joseph", "Oropuna")  

#Drainage level
Caroni<-c("Tunapuna", "Lopinot", "Guanapo", "Aripo", "Cuara", "Tranquille", "Caroni", "Arima", "Arouca", "Mausica", "Tacarigua", "El Cedro", "Ceniza", "Naranjo", "Maracas", "Caroni Lower Dyke", "Caura", "Manacal", "Saint Joseph", "Oropuna")

Oropuche<-c("Quare", "Hollis", "Oropuche", "Turure", "La Seiva", "Rio Barro", "Rio Grande", "La Selva", "Valencia", "Matura", "Hollis")

#rivers not already sorted into north or south slope
#unknown.rivers<-subset(river.list, (!(river.list$Population %in% North | river.list$Population %in% South| river.list$Population %in% Northeastern| river.list$Population %in% Other))) 

#For mapping watershed sorting STILL NEED(Maracas, Saint Joseph, Manacal, Tranquille, Mausica, Blanadra)
map_other<-c("Pilote", "Lizard", "Guapo", "Coffee", "Silverstream", "Arena", "Vance", "Mayaro", "Cunupia Lower Dyke", "Manzanilla", "Visigney", "Tompire", "Ricon", "Mission", "Toco", "Mausica", "Maracas")
map_aripo<-c("Aripo", "Ceniza", "Naranjo", "Valencia")
map_quare<-c("Quare","Hollis")
map_tunapuna<-c("Tunapuna", "Caroni Lower Dyke")
map_tacarigua<-c("Tacarigua", "Caura", "Manacal")
map_arouca<-c("Arouca", "Lopinot", "Oropuna")
map_guanapo<-c("Guanapo")
map_arima<-c("Arima")
map_elcedro<-c("El Cedro")
map_turure<-c("Turure")
map_oropuche<-c("Oropuche", "Matura", "La Seiva", "Rio Barre", "Rio Grande", "La Selva")
map_yarra<-c("Yarra")
map_marianne<-c("Marianne", "Brasso Seco")
map_paria<-c("Paria", "Madamas")
map_damier<-c("Damier")
data$pop_map<-case_when(
  data$Population %in% map_aripo ~ "Aripo",
  data$Population %in% map_quare ~ "Quare",
  data$Population %in% map_tacarigua ~ "Tacarigua",
  data$Population %in% map_tunapuna ~ "Tunapuna",
  data$Population %in% map_arouca ~ "Arouca",
  data$Population %in% map_guanapo ~ "Guanapo",
  data$Population %in% map_arima ~ "Arima",
  data$Population %in% map_elcedro ~ "El Cedro",
  data$Population %in% map_turure ~ "Turure",
  data$Population %in% map_oropuche ~ "Oropuche",
  data$Population %in% map_yarra ~ "Yarra",
  data$Population %in% map_marianne ~ "Marianne",
  data$Population %in% map_paria ~ "Paria",
  data$Population %in% map_damier ~ "Damier",
  data$Population %in% map_other ~ "Other"
)
```



```{r Slope, Drainage, and TraitID columns, eval=FALSE}

#new.data<-data #!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! don't forget this you dummy

new.data$Slope<-  case_when(
  new.data$Population %in% South ~ "South",
  new.data$Population %in% North ~ "North",
  new.data$Population %in% Other ~ "Other",
  new.data$Population %in% Northeastern ~ "Northeastern"
) #labels all single populations as North or South Slope, and other for identified rivers not in this dichotomy

new.data$Drainage<- case_when(
  new.data$Population %in% Caroni ~ "Caroni",
  new.data$Population %in% Oropuche ~ "Oropuche",
  new.data$Population %in% North ~ "Northern",
  new.data$Population %in% Other ~ "Other",
  new.data$Population %in% Northeastern ~ "Other",
)  #Labels all single populations as a specific drainage


#coding the trait column, one ID for each trait measured in each sex group in each study
new.data<-new.data %>% 
  arrange(StudyID, StudyType, Trait, Sex, Predation)
sex.length<-rle(as.character(paste(new.data$Trait,new.data$Sex,new.data$StudyType)))  
new.data$TraitID<-rep(seq_along(sex.length$lengths), sex.length$lengths)
#also recoding studyID because of excluded studies making the numbers janky
study.length<-rle(as.character(paste(new.data$Study)))
new.data$StudyID<-rep(seq_along(study.length$lengths), study.length$lengths)

#fixes some kingsolver trait mis-sorting (March 10th meeting), final spreadsheet done manually (july 21)
#new.data$Kingsolver_traits<-ifelse(new.data$TraitID %in% c(41:42,62:79,81:82,89:91,119:121,295:297,301:303,554), "Physiology",ifelse(new.data$TraitID %in% c(27:28,33:34,377,472), "Other_life_history",ifelse(new.data$TraitID %in% c(471,473), "Other_morphology", new.data$Kingsolver_traits)))

#changing intro sites, all turure and damier, El cedro LP is intro, El Cedro HP is natural, Arima after 2001 HP is intro
new.data$Poptype_broad<-ifelse(new.data$Population %in% c("Damier", "Turure"), "Introduction",
                               ifelse((new.data$Population %in% c("El Cedro") & new.data$Predation=="Low"), "Introduction", 
                                      ifelse((new.data$Population %in% c("El Cedro") & new.data$Predation=="Low"), "Introduction", 
                                             ifelse((new.data$Population %in% c("Arima") & new.data$Predation=="High" & new.data$Collection_start>2001), "Introduction", new.data$Poptype))))

```


testing a slope column that indicates if within a trait populations are all from the southern slope (south), are all from northern slope (north), or are from both (mixed). Also need to know how many of each we have
```{r overall slope column}
slope.data<-new.data %>% 
  group_by(TraitID) %>% 
  mutate(unique_slopes = n_distinct(Slope))
slope.data$Slope.type<-ifelse(slope.data$unique_slopes==1, slope.data$Slope, "Mixed")

summary(as.factor(slope.data$unique_slopes))
summary(as.factor(slope.data$Slope.type))

x<-slope.data[!duplicated(slope.data$TraitID),]
summary(as.factor(x$unique_slopes))
summary(as.factor(x$Slope.type))

#write_csv(slope.data, path = "~/Documents/Masters/R scripts/Meta Analysis/Data/MetaData.csv") #written July 21st
```


### yeardata formatting (no longer needed, integrated into metaData)
```{r years, eval=FALSE}
year<-new.data %>% 
  mutate(duration =  (as.numeric(str_sub(CollectionYear, -4))) - (as.numeric(str_sub(CollectionYear, start = 0, 4)))) %>% 
  mutate(YearStart = as.numeric(str_sub(CollectionYear, start = 0, 4))) %>% 
  filter(duration < 2) #27 studies under 2 years, 4 studies with a higher range (Aug 6th)

#year$YearEnd<-as.numeric(str_sub(year$CollectionYear, -4)) #if we want the last year of collection


traitData<-left_join(r.squared, year, by = "TraitID")
traitData<-traitData[!duplicated(traitData$TraitID),]
```

### Mapping miscellaneous 

```{r mapping prep}
Coordinates <- read_csv("~/Documents/Projects/R scripts/Meta Analysis/Data/Map/most_coordinates.csv")
Coordinates<-Coordinates %>% 
  select(1,2,4,7,17:19,21:23,25,29:30,33:35) %>% 
  arrange(TraitID)

# I'm just being lazy and copying this code again instead of formatting for a join
map_other<-c("Pilote", "Lizard", "Guapo", "Coffee", "Silverstream", "Arena", "Vance", "Mayaro", "Cunupia Lower Dyke", "Manzanilla", "Visigney", "Tompire", "Ricon", "Mission", "Toco", "Mausica", "Maracas")
map_aripo<-c("Aripo", "Ceniza", "Naranjo", "Valencia")
map_quare<-c("Quare","Hollis")
map_tunapuna<-c("Tunapuna", "Caroni Lower Dyke")
map_tacarigua<-c("Tacarigua", "Caura", "Manacal")
map_arouca<-c("Arouca", "Lopinot", "Oropuna")
map_guanapo<-c("Guanapo")
map_arima<-c("Arima")
map_elcedro<-c("El Cedro")
map_turure<-c("Turure")
map_oropuche<-c("Oropuche", "Matura", "La Seiva", "Rio Barre", "Rio Grande", "La Selva")
map_yarra<-c("Yarra")
map_marianne<-c("Marianne", "Brasso Seco")
map_paria<-c("Paria", "Madamas")
map_damier<-c("Damier")
Coordinates$pop_map<-case_when(
  Coordinates$Population %in% map_aripo ~ "Aripo",
  Coordinates$Population %in% map_quare ~ "Quare",
  Coordinates$Population %in% map_tacarigua ~ "Tacarigua",
  Coordinates$Population %in% map_tunapuna ~ "Tunapuna",
  Coordinates$Population %in% map_arouca ~ "Arouca",
  Coordinates$Population %in% map_guanapo ~ "Guanapo",
  Coordinates$Population %in% map_arima ~ "Arima",
  Coordinates$Population %in% map_elcedro ~ "El Cedro",
  Coordinates$Population %in% map_turure ~ "Turure",
  Coordinates$Population %in% map_oropuche ~ "Oropuche",
  Coordinates$Population %in% map_yarra ~ "Yarra",
  Coordinates$Population %in% map_marianne ~ "Marianne",
  Coordinates$Population %in% map_paria ~ "Paria",
  Coordinates$Population %in% map_damier ~ "Damier",
  Coordinates$Population %in% map_other ~ "Other"
)
# this codes the poptype_broad column to indicate our system knowledge that certain watersheds were not colonized by guppies until researchers' experiments
Coordinates$Poptype_broad<-ifelse(Coordinates$Population %in% c("Damier", "Turure"), "Introduction",
                               ifelse((Coordinates$Population %in% c("El Cedro") & Coordinates$Predation=="Low"), "Introduction", 
                                      ifelse((Coordinates$Population %in% c("El Cedro") & Coordinates$Predation=="Low"), "Introduction", 
                                             ifelse((Coordinates$Population %in% c("Arima") & Coordinates$Predation=="High" & Coordinates$Collection_start>2001), "Introduction", Coordinates$Poptype))))

#write_csv(Coordinates, path = "~/Documents/Projects/R scripts/Meta Analysis/Data/Map/site_type.csv")

```

### Follow Up on reviewer comment, R2 permutations

```{r functions}

dat<-FALSE 

i <- 1

anova_loop.resample <- function(dat=dat){
  
  trait <- as.factor(dat$TraitID)

#output.perm <- array(dim=c(1000, 3, (length(unique(dat$TraitID)))))
output.perm <- array(dim=c(100, 3, (length(unique(dat$TraitID)))))

  
  for (i in 1: length(levels(trait))){
    dat.temp <- dat[dat$TraitID==levels(trait)[i],]
    print(i)
    traitID <- levels(trait)[i]
    
    #resampling permutations

output.permj <- matrix (ncol=3, nrow=(100))
colnames(output.permj) <- c("TraitID", "R2.RE", "RadjRE")

for (j in 1:100){

  dat.temp$PredationRESAMPLE <- sample(dat.temp$Predation, replace = FALSE)
  
  aov.modRE   <- aov(MeanValue~PredationRESAMPLE,data=dat.temp)
  anova.modRE <- stats:::anova.lm(aov.modRE)
  mod.ssRE    <- anova.modRE$"Sum Sq"
  mod.pesRE   <- mod.ssRE/(mod.ssRE+mod.ssRE[length(mod.ssRE)])          # calculate pes
  
  rRE <- summary.lm(aov.modRE)                                     # summary of linear model
  
  R2.RE     <- round(rRE$"r.squared",3)                              # R squared
  R_adjRE <- round(rRE$"adj.r.squared",3)
  

  #output.RE[j,] <- c(R,R_adj,partial_eta,F_value,p_value, traitID, justr, R2.RE, R_adjRE) 
  output.permj[j,] <- c(traitID, R2.RE, R_adjRE)
}
 output.perm[,,i] <- output.permj   

  }
output.perm<-as.data.frame(output.perm)
colnames(output.perm)<- rep(c("TraitID", "R2", "R2adj"),len = (ncol(output.perm))) #rename the columns as data type

#reformats the output as a nice df
output.perm<-pivot_longer(output.perm, cols = 1:3, names_to = 'type') 
output.perm<-pivot_wider(output.perm, names_from = type, values_from = value)

resampleR2<-data.frame(TraitID = (unlist(output.perm$TraitID)), R2 = (unlist(output.perm$R2)), R2adj = (unlist(output.perm$R2adj))) %>%
  mutate(R2 = as.numeric(R2), TraitID = as.numeric(TraitID), R2adj = as.numeric(R2adj))
return(resampleR2)

}

full_output<-function(dat = dat, resampleDat = resampleDat){
    
  #calculate the mean R2 per trait
R2means<-resampleDat %>% 
  group_by(TraitID) %>% 
  summarize(R2mean = mean(R2))

  #create output df with all R2 data
rDat <- left_join(dat, R2means, by = 'TraitID') #merge by TraitID (same col name)
    
  #run model with original and mean perm R2
ex.mod <- lm(R.2 ~ R2mean, data=rDat)
resid <- residuals(ex.mod)
rDat$resid <- resid
  
return(rDat)
}

```




```{r loading all dfs}
# load in all data tables as unique outputs
#The trait filter groups are in Tallies.rmd under ## Subsets for Among and Within, commented the corresponding df for reference
output_all<-read.csv(paste(wd,'/Data/TraitR2_among.csv',sep="")) #new.data
output_south<-read.csv(paste(wd,'/Data/TraitR2_South.csv',sep="")) #traits_s
output_caroni<-read.csv(paste(wd,'/Data/TraitR2_Caroni.csv',sep=""))#traits_d
output_south_drainages<-read.csv(paste(wd,'/Data/TraitR2_Among_Drainage.csv',sep="")) #traits_d_all
output_intros<-read.csv(paste(wd,'/Data/TraitR2_intro.csv',sep="")) #traits_i
output_intros_broad<-read.csv(paste(wd,'/Data/TraitR2_intro_broad.csv',sep="")) #traits_i_broad

```



```{r creating outputs}
#running the permutations for each group
resample_all<-anova_loop.resample(dat = new.data)
resample_south<-anova_loop.resample(dat = traits_s)
resample_caroni<-anova_loop.resample(dat = traits_d)
resample_south_drainages<-anova_loop.resample(dat = traits_d_all)
resample_intros<-anova_loop.resample(dat = traits_i)
resample_intros_broad<-anova_loop.resample(dat = traits_i_broad)

#creating nice outputs for each
Rdat_all<-full_output(dat = output_all, resampleDat = resample_all)
Rdat_south<-full_output(dat = output_south, resampleDat = resample_south)
Rdat_caroni<-full_output(dat = output_caroni, resampleDat = resample_caroni)
Rdat_south_drainages<-full_output(dat = output_south_drainages, resampleDat = resample_south_drainages)
Rdat_intros<-full_output(dat = output_intros, resampleDat = resample_intros)
#Rdat_intros_broad<-full_output(dat = output_intros_broad, resampleDat = resample_intros_broad) # error

#writing outputs to file, R2mean is the mean perm R2 per trait, resid is the model residuals
#write results to a csv file for further analyses DONT FORGET TO MOVE AND OVERWRITE saves inside repo not data folder

#this is the overall R2, all studies and traits included
Rdat_all<-full_output(dat = output_all, resampleDat = resample_all)
#write.table(Rdat_all, file = "Resample_among.csv",row.names=FALSE,col.names=TRUE, sep=",") 

# for the south only (Slope Q)
Rdat_south<-full_output(dat = output_south, resampleDat = resample_south)
#write.table(Rdat_south, file = "Resample_South.csv",row.names=FALSE,col.names=TRUE, sep=",") 

# for the caroni only (drainage Q)
Rdat_caroni<-full_output(dat = output_caroni, resampleDat = resample_caroni)
#write.table(Rdat_caroni, file = "Resample_Caroni.csv",row.names=FALSE,col.names=TRUE, sep=",") 

# for the south only (drainage Q, northern pops excluded because they arent in caroni or oropuche drainages)
Rdat_south_drainages<-full_output(dat = output_south_drainages, resampleDat = resample_south_drainages)
#write.table(Rdat_south_drainages, file = "Resample_Among_Drainage.csv",row.names=FALSE,col.names=TRUE, sep=",") 

# anova for the natural pops only (drainage Q)
Rdat_intros<-full_output(dat = output_intros, resampleDat = resample_intros)
#write.table(Rdat_intros, file = "Resample_intro.csv",row.names=FALSE,col.names=TRUE, sep=",")

```


```{r intros_broad}
# for the natural pops only (drainage Q, broad specification)
#issues with intros_broad (trait 303 creates NaN response in model), drop that trait
  #calculate the mean R2 per trait
means_i<-resample_intros_broad %>% 
  filter(!TraitID ==303) %>% 
  group_by(TraitID) %>% 
  summarize(R2mean = mean(R2))

  #create output df with all R2 data
Rdat_intros_broad <- left_join((output_intros_broad %>% 
                            filter(!TraitID == 303)), means_i, by = 'TraitID') #merge by TraitID (same col name)
    
  #run model with original and mean perm R2
ex.mod <- lm(R.2 ~ R2mean, data=Rdat_intros_broad)
resid <- residuals(ex.mod)
Rdat_intros_broad$resid <- resid 

#write.table(Rdat_intros_broad, file = "Resample_intro_broad.csv",row.names=FALSE,col.names=TRUE, sep=",") 

```




```{r visualization}
# for all the traits
plot(Rdat_all$R2mean, Rdat_all$R.2)
hist(Rdat_all$R2mean)
hist(Rdat_all$resid)

plot_a <- ggplot(Rdat_all, aes(R2mean, R.2))

plot_a + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R² (all)") + xlab("Mean R² (all) expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

# for south

plot(Rdat_south$R2mean, Rdat_south$R.2)
hist(Rdat_south$R2mean)
hist(Rdat_south$resid)

plot_s <- ggplot(Rdat_south, aes(R2mean, R.2))

plot_s + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R² (south)") + xlab("Mean R² (south) expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

# for south_drainages

plot(Rdat_south_drainages$R2mean, Rdat_south_drainages$R.2)
hist(Rdat_south_drainages$R2mean)
hist(Rdat_south_drainages$resid)

plot_sd <- ggplot(Rdat_south_drainages, aes(R2mean, R.2))

plot_sd + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R² (south drainages)") + xlab("Mean R² (south drainages) expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

# for caroni
plot(Rdat_caroni$R2mean, Rdat_caroni$R.2)
hist(Rdat_caroni$R2mean)
hist(Rdat_caroni$resid)

plot_c <- ggplot(Rdat_caroni, aes(R2mean, R.2))

plot_c + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R² (caroni)") + xlab("Mean R² (caroni) expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

# for intros
plot(Rdat_intros$R2mean, Rdat_intros$R.2)
hist(Rdat_intros$R2mean)
hist(Rdat_intros$resid)

plot_i <- ggplot(Rdat_intros, aes(R2mean, R.2))

plot_i + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R² (intros)") + xlab("Mean R² (intros) expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

```

FYI Andrew's response to the email about the sample size plot: "Possibly, yes, but you would want to fit a function to this to see if there really is a decline. I am not sure that is the case. I suppose you could do a linear fit (perhaps no decrease) and a fit to the upper quantile to show that maximum values tend to be at the lowest sample sizes. Regardless, the plot is cool and I would like to include it somehow."
```{r sample size}
all_dat<-left_join(output_all, new.data[,c(1,3,20,37)], by = 'TraitID')
### find outliers
isnt_out_z <- function(x, thres = 3, na.rm = TRUE) {
  abs(x - mean(x, na.rm = na.rm)) <= thres * sd(x, na.rm = na.rm)
}
all_dat$outlier<-isnt_out_z(all_dat$Number)

#run simple models
all_mod<-lm(R.2~Number, (all_dat  %>% 
           filter(!is.na(Number))))
filter_mod<-lm(R.2~Number, (all_dat %>% 
                              filter(!is.na(Number) & outlier==TRUE)))

summary(all_mod)
summary(filter_mod)

## what was I doing? making a proper model?
test_dat<-all_dat %>% 
  filter(!is.na(Number))
test_mod<-lmer(R.2~Number + (1|StudyID), all_dat)
summary(test_mod)

```




```{r repeating R2}
#similar R2 in the plots??
sus<-left_join(new.data, output_all, by = 'TraitID') %>% 
  select(1, 20:21,37, 46:50)

sus %>% 
  group_by(StudyID) %>% 
  summarize(n = n_distinct(R.2), n_trait = n_distinct(TraitID), diff = (n_trait - n)) %>% 
  filter(diff>0)

sus2<-sus %>% 
  filter(StudyID %in% c(4,9,10,15,18,23,32,37,40,47)) %>% #studies with repeating R2 
  group_by(TraitID) %>% 
  summarize(StudyID = median(StudyID), R2 = median(R.2))


#sample size plot
N_plot<-ggplot(all_dat, aes(Number, R.2)) + geom_point() +theme_bw() + theme(axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        plot.title = element_text(size = 16)) +labs(x = "Sample Size", y = "R2")
ggsave("N_plot.pdf")

### visualizations from meeting
#ggplot((all_dat %>% 
         #filter(!(StudyID == 22|StudyID ==20))), aes(Number, R.2, color = (Published))) + geom_point()+ scale_color_gradient(low="red", high="yellow")
```



