---
title: "Slopes & Drainages"
output: html_notebook
---


```{r}
library(tidyverse)
data<- read.csv('YOUR_DATA.csv', header=TRUE, sep=",") %>% 
  rename( Mean = MeanValue, Treatment = Predation)

data<- read.csv('../Data/data.csv', header=TRUE, sep=",") %>% 
    rename( Mean = MeanValue, Treatment = Predation)
```

Add in a column noting if north or south slope watersheds, and one for drainage
North: 
Yarra, Marianne, Paria, Madamas, Mission (North but over near Tompire?), Tompire, Damier

South: 
Caroni Drainage (Tunapuna, Lopinot, Guanapo, Aripo, Cuara, Tranquille, Mausica, Tacarigua, Arouca, El Cedro, Cuara)
Oropuche Drainage (Quare, Oropuche, Turure, Matura - but northeast, La Seiva)

Questionable
Tompire: west coast potentially 'north' but not connected to others
Arena: Near the arena reservoir some place? thoroughly out of the mountains
Naranjo: north but kinda like tompire
Tumbason:south, drainage? not included anymore
Hollis: South? not included anymore
Matura: in oropuche drainage but far northeast
Silverstream: South West (from jackie)
Vance: South West (from jackie, I think it means not in oropuche or caroni, potentially out of the south slope too)


```{r drainage slope}
data$Population<-gsub("ElCedro1", "El Cedro", data$Population)
data$Population<-gsub("ElCedro2", "El Cedro", data$Population)
data$Population<-gsub("Lower Guapo 1", "Guapo", data$Population)
data$Population<-gsub("Lower Guapo 2", "Guapo", data$Population)
data$Population<-gsub("laseiva", "La Seiva", data$Population)

data$Population<-as.factor(data$Population)
river.list<-data %>% 
  filter(PopulationType=="Single") %>% 
  select(Population) %>% 
  unique()

North<-c("Yarra", "Marianne", "Paria", "Madamas", "Damier") #"Mission", "Tompire", "Naranjo"
South<-c("Arima", "Tunapuna", "Lopinot", "Guanapo", "Aripo", "Cuara", "Tranquille", "Quare", "Oropuche", "Turure", "Matura", "Caroni", "Arouca", "Mausica", "Tacarigua", "El Cedro")

@Drainage level
Caroni<-c("Tunapuna", "Lopinot", "Guanapo", "Aripo", "Cuara", "Tranquille", "Caroni", "Arima", "Arouca", "Mausica", "Tacarigua", "El Cedro", "Caroni Lower Dyke")

Oropuche<-c("Quare", "Oropuche", "Turure", "La Seiva") #"Matura"

unknown.rivers<-subset(river.list, (!(river.list$Population %in% North | river.list$Population %in% South)))
print(unknown.rivers)
```

We still need to locate these and identify which drainage if possible as well (x24): 
Balandra			
Brasso Seco   (North? the village is north slope)	
Cenzina			
Rio Barro			
Rio Grande			
Tompire		(Far north east?)	
Ceniza			
QT        (quare trib?)		
Ricon			
La Selva			
Cunupia Lower Dyke			
Caura			
Coffee			
Guapo			
Lizard			
Manzanilla			
Pilote			
Toco			
Lower Trib 1			
Lower Trib 2			
Visigney			
Maracas		(North based on the beach/town?)	
Vance			
Manacal			
Valencia    (south maybe oropuche (based on town))			


```{r columns}

data$Slope<-  case_when(
  data$Population %in% South ~ "South",
  data$Population %in% North ~ "North" #labels all single populations as North or South Slope
) 
data$Drainage<- case_when(
  data$Population %in% Caroni ~ "Caroni",
  data$Population %in% Oropuche ~ "Oropuche",
  data$Population %in% North ~ "Northern" #Labels all single populations as a specific drainage
)
```
