---
title: "Slopes & Drainages"
output: html_notebook
---


```{r read in}
library(tidyverse)
library(ggplot2)
#setwd("<YOURPATHWAY>")
wd <- getwd()
new.data<- read.csv(paste(wd,'/Data/MetaData.csv',sep=""), header=TRUE, sep=",") #MetaData saved feb 11th
data<- read.csv(paste(wd,'/Data/Data.csv',sep=""), header=TRUE, sep=",")%>% #Data.csv from .xlsx feb 25th
  select(StudyID:Comments) 
r.squared<- read.csv(paste(wd,'/Data/TraitR2_among.csv',sep=""), header=TRUE, sep=",") #R2 generated feb 11th 
```

###Misc Notes on populations
North: 
Yarra, Marianne, Paria, Madamas, Damier, Brasso Seco, Balandra *Tompire*, *Ricon*, *Mission*, *Toco* (*northeast?*)

South: 
Caroni Drainage (Tunapuna, Lopinot, Guanapo, Aripo, Cuara, Tranquille, Caroni, Arima, Arouca, Mausica, Tacarigua, El Cedro, Ceniza, Naranjo, Maracas, Caroni Lower Dyke, Cuara, Manacal)
Oropuche Drainage (Quare, Oropuche, Turure, La Seiva, Rio Barro, Rio Grande, La Selva, Valencia, Matura)

*Questionable/Notes*
Madamas: Potentially a northern drainage, it's right beside the Paria to the east
Naranjo: upper aripo
Tumbason:south, drainage? not included anymore
Hollis: South? not included anymore
Tompire: west coast usually counted as northern range but not connected to others 
Mission: North but over near Tompire
Ricon: north but east and not connected to a drainage
Toco: north east like tompire
Cunupia Lower Dyke: near caroni but listed as distinct in 23, has gps in supplementary mat.	
Manzanilla:	No gps, listed as separate drainage
Matura: inlcuded for now in oropuche but some say it's not because its far east and north a bit
Cenzina changed to Ceniza (typo confirmed in paper)

*Southern trinidad outside northern mountain range*

Southern trinidad outside northern mountain range
Pilote and lower trib 1 &2 are a part of a Mayaro drainage, in southern trinidad)
Lizard
Guapo
Coffee
Visigney
Silverstream
Arena: Near the arena reservoir some place? thoroughly out of the mountains
Vance: South West (from jackie outside the northern range, also in paper 23, but not listed in supplementary material as a site??)

### Formatting columns
```{r population column}
data$temp<-NA #temp columns
data$pop<-data$PopulationRaw

names<-c("Arima","Arena","Arouca","Aripo", "Quare", "Yarra", "Tompire", "Marianne", "Paria", "El Cedro", "Madamas", "Oropuche", "Guanapo", "La Selva", "Turure", "Damier", "Tacarigua", "Mausica", "Manacal", "Ceniza", "Caroni", "Valencia", "Oropuna", "Tunapuna", "Lopinot", "Guapo", "Vance", "Visigney", "Pilote", "Lizard", "Silverstream", "Matura","Cunupia", "Coffee", "Maracas", "Caura", "Toco", "Manzanilla", "Mission", "Ricon", "Balandra", "Brasso Seco", "Rio Grande", "Rio Barro", "Caroni Lower Dyke", "Cunupia Lower Dyke")
for (i in names){ data<-mutate(data, pop = (ifelse(grepl(i, PopulationRaw), i,  temp)))
data$temp<-ifelse(data$pop==i, i, data$temp)} #replaces variations of populations with the root, ie Aripo 4 -> Aripo

data$pop<-ifelse(is.na(data$pop),data$PopulationRaw, data$pop) #fill NAs with the raw population data

#fix typos

data$pop<-gsub("ElCedro1", "El Cedro", data$pop)
data$pop<-gsub("ElCedro2", "El Cedro", data$pop)
data$pop<-gsub("Lower Guapo 1", "Guapo", data$pop)
data$pop<-gsub("Lower Guapo 2", "Guapo", data$pop)
data$pop<-gsub("Laseiva", "La Seiva", data$pop)
data$pop<-gsub("Cenzina", "Ceniza", data$pop)
data$pop<-gsub("Lower Trib 1", "Mayaro", data$pop)
data$pop<-gsub("Lower Trib 2", "Mayaro", data$pop)
data$pop<-gsub("Tompire 2", "Tompire", data$pop)
data$pop<-gsub("guanapo", "Guanapo", data$pop)
data$pop<-gsub("Lower La Seiva", "La Seiva", data$pop)
data$pop<-gsub("Upper turure", "Turure", data$pop)
data$pop<-gsub("Tacargua", "Tacarigua", data$pop)
data$pop<-gsub("Oropouche", "Oropuche", data$pop)
data$pop<-gsub("St Joseph", "Saint Joseph", data$pop)
data$pop<-gsub("Tac/Tig", "Tacarigua", data$pop) # Need to confirm this!!!!
#data$pop<-ifelse(grepl("Dyke", data$PopulationRaw), data$PopulationRaw, data$pop)

summary(as.factor(data$pop))
data$Population<-data$pop
data<-data %>% 
  select(StudyID:Comments)
```



```{r typos and river lists, eval=FALSE}
#fix typos not previously found
data$TraitType2<-gsub("Life-history", "Life history", data$TraitType2)
data$PopulationRaw<-gsub("Oropouche", "Oropuche", data$PopulationRaw)

data$Population<-as.factor(data$Population)
river.list<-data %>% 
  select(Population) %>% 
  unique()

North<-c("Yarra", "Marianne", "Paria", "Madamas", "Damier", "Brasso Seco", "Balandra") #I think we should keep these together for drainage level, logistically it'll be easier and its how others do it (paper 2 is an example)

Northeastern<- c("Tompire", "Ricon", "Mission", "Toco")

Other<-c("Pilote", "Lizard", "Guapo", "Coffee", "Silverstream", "Arena", "Vance", "Mayaro", "Cunupia Lower Dyke", "Manzanilla", "Visigney") #mostly southern range, jessica stephenson paper, but some others

South<-c("Tunapuna", "Lopinot", "Guanapo", "Aripo", "Caura", "Tranquille", "Caroni", "Arima", "Arouca", "Mausica", "Tacarigua", "El Cedro", "Ceniza", "Naranjo", "Maracas", "Caroni Lower Dyke", "Manacal", "Quare","Hollis", "Oropuche", "Turure", "La Seiva", "Rio Barro", "Rio Grande", "La Selva", "Valencia", "Matura", "Hollis", "Saint Joseph", "Oropuna")  

#Drainage level
Caroni<-c("Tunapuna", "Lopinot", "Guanapo", "Aripo", "Cuara", "Tranquille", "Caroni", "Arima", "Arouca", "Mausica", "Tacarigua", "El Cedro", "Ceniza", "Naranjo", "Maracas", "Caroni Lower Dyke", "Caura", "Manacal", "Saint Joseph", "Oropuna")

Oropuche<-c("Quare", "Hollis", "Oropuche", "Turure", "La Seiva", "Rio Barro", "Rio Grande", "La Selva", "Valencia", "Matura", "Hollis")

#rivers not already sorted into north or south slope
#unknown.rivers<-subset(river.list, (!(river.list$Population %in% North | river.list$Population %in% South| river.list$Population %in% Northeastern| river.list$Population %in% Other))) 

#For mapping watershed sorting STILL NEED(Maracas, Saint Joseph, Manacal, Tranquille, Mausica, Blanadra)
map_other<-c("Pilote", "Lizard", "Guapo", "Coffee", "Silverstream", "Arena", "Vance", "Mayaro", "Cunupia Lower Dyke", "Manzanilla", "Visigney", "Tompire", "Ricon", "Mission", "Toco", "Mausica", "Maracas")
map_aripo<-c("Aripo", "Ceniza", "Naranjo", "Valencia")
map_quare<-c("Quare","Hollis")
map_tunapuna<-c("Tunapuna", "Caroni Lower Dyke")
map_tacarigua<-c("Tacarigua", "Caura", "Manacal")
map_arouca<-c("Arouca", "Lopinot", "Oropuna")
map_guanapo<-c("Guanapo")
map_arima<-c("Arima")
map_elcedro<-c("El Cedro")
map_turure<-c("Turure")
map_oropuche<-c("Oropuche", "Matura", "La Seiva", "Rio Barre", "Rio Grande", "La Selva")
map_yarra<-c("Yarra")
map_marianne<-c("Marianne", "Brasso Seco")
map_paria<-c("Paria", "Madamas")
map_damier<-c("Damier")
data$pop_map<-case_when(
  data$Population %in% map_aripo ~ "Aripo",
  data$Population %in% map_quare ~ "Quare",
  data$Population %in% map_tacarigua ~ "Tacarigua",
  data$Population %in% map_tunapuna ~ "Tunapuna",
  data$Population %in% map_arouca ~ "Arouca",
  data$Population %in% map_guanapo ~ "Guanapo",
  data$Population %in% map_arima ~ "Arima",
  data$Population %in% map_elcedro ~ "El Cedro",
  data$Population %in% map_turure ~ "Turure",
  data$Population %in% map_oropuche ~ "Oropuche",
  data$Population %in% map_yarra ~ "Yarra",
  data$Population %in% map_marianne ~ "Marianne",
  data$Population %in% map_paria ~ "Paria",
  data$Population %in% map_damier ~ "Damier",
  data$Population %in% map_other ~ "Other"
)
```



```{r Slope, Drainage, and TraitID columns, eval=FALSE}

#new.data<-data #!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!! don't forget this you dummy

new.data$Slope<-  case_when(
  new.data$Population %in% South ~ "South",
  new.data$Population %in% North ~ "North",
  new.data$Population %in% Other ~ "Other",
  new.data$Population %in% Northeastern ~ "Northeastern"
) #labels all single populations as North or South Slope, and other for identified rivers not in this dichotomy

new.data$Drainage<- case_when(
  new.data$Population %in% Caroni ~ "Caroni",
  new.data$Population %in% Oropuche ~ "Oropuche",
  new.data$Population %in% North ~ "Northern",
  new.data$Population %in% Other ~ "Other",
  new.data$Population %in% Northeastern ~ "Other",
)  #Labels all single populations as a specific drainage


#coding the trait column, one ID for each trait measured in each sex group in each study
new.data<-new.data %>% 
  arrange(StudyID, StudyType, Trait, Sex, Predation)
sex.length<-rle(as.character(paste(new.data$Trait,new.data$Sex,new.data$StudyType)))  
new.data$TraitID<-rep(seq_along(sex.length$lengths), sex.length$lengths)
#also recoding studyID because of excluded studies making the numbers janky
study.length<-rle(as.character(paste(new.data$Study)))
new.data$StudyID<-rep(seq_along(study.length$lengths), study.length$lengths)

#fixes some kingsolver trait mis-sorting (March 10th meeting), final spreadsheet done manually (july 21)
#new.data$Kingsolver_traits<-ifelse(new.data$TraitID %in% c(41:42,62:79,81:82,89:91,119:121,295:297,301:303,554), "Physiology",ifelse(new.data$TraitID %in% c(27:28,33:34,377,472), "Other_life_history",ifelse(new.data$TraitID %in% c(471,473), "Other_morphology", new.data$Kingsolver_traits)))

#changing intro sites, all turure and damier, El cedro LP is intro, El Cedro HP is natural, Arima after 2001 HP is intro
new.data$Poptype_broad<-ifelse(new.data$Population %in% c("Damier", "Turure"), "Introduction",
                               ifelse((new.data$Population %in% c("El Cedro") & new.data$Predation=="Low"), "Introduction", 
                                      ifelse((new.data$Population %in% c("El Cedro") & new.data$Predation=="Low"), "Introduction", 
                                             ifelse((new.data$Population %in% c("Arima") & new.data$Predation=="High" & new.data$Collection_start>2001), "Introduction", new.data$Poptype))))

```


testing a slope column that indicates if within a trait populations are all from the southern slope (south), are all from northern slope (north), or are from both (mixed). Also need to know how many of each we have
```{r overall slope column}
slope.data<-new.data %>% 
  group_by(TraitID) %>% 
  mutate(unique_slopes = n_distinct(Slope))
slope.data$Slope.type<-ifelse(slope.data$unique_slopes==1, slope.data$Slope, "Mixed")

summary(as.factor(slope.data$unique_slopes))
summary(as.factor(slope.data$Slope.type))

x<-slope.data[!duplicated(slope.data$TraitID),]
summary(as.factor(x$unique_slopes))
summary(as.factor(x$Slope.type))

#write_csv(slope.data, path = "~/Documents/Masters/R scripts/Meta Analysis/Data/MetaData.csv") #written July 21st
```


### yeardata formatting (no longer needed, integrated into metaData)
```{r years, eval=FALSE}
year<-new.data %>% 
  mutate(duration =  (as.numeric(str_sub(CollectionYear, -4))) - (as.numeric(str_sub(CollectionYear, start = 0, 4)))) %>% 
  mutate(YearStart = as.numeric(str_sub(CollectionYear, start = 0, 4))) %>% 
  filter(duration < 2) #27 studies under 2 years, 4 studies with a higher range (Aug 6th)

#year$YearEnd<-as.numeric(str_sub(year$CollectionYear, -4)) #if we want the last year of collection


traitData<-left_join(r.squared, year, by = "TraitID")
traitData<-traitData[!duplicated(traitData$TraitID),]
```

### Mapping miscellaneous ###


```{r mapping prep}
Coordinates <- read_csv("~/Documents/Projects/R scripts/Meta Analysis/Data/Map/most_coordinates.csv")
Coordinates<-Coordinates %>% 
  select(1,2,4,7,17:19,21:23,25,29:30,33:35) %>% 
  arrange(TraitID)

# I'm just being lazy and copying this code again instead of formatting for a join
map_other<-c("Pilote", "Lizard", "Guapo", "Coffee", "Silverstream", "Arena", "Vance", "Mayaro", "Cunupia Lower Dyke", "Manzanilla", "Visigney", "Tompire", "Ricon", "Mission", "Toco", "Mausica", "Maracas")
map_aripo<-c("Aripo", "Ceniza", "Naranjo", "Valencia")
map_quare<-c("Quare","Hollis")
map_tunapuna<-c("Tunapuna", "Caroni Lower Dyke")
map_tacarigua<-c("Tacarigua", "Caura", "Manacal")
map_arouca<-c("Arouca", "Lopinot", "Oropuna")
map_guanapo<-c("Guanapo")
map_arima<-c("Arima")
map_elcedro<-c("El Cedro")
map_turure<-c("Turure")
map_oropuche<-c("Oropuche", "Matura", "La Seiva", "Rio Barre", "Rio Grande", "La Selva")
map_yarra<-c("Yarra")
map_marianne<-c("Marianne", "Brasso Seco")
map_paria<-c("Paria", "Madamas")
map_damier<-c("Damier")
Coordinates$pop_map<-case_when(
  Coordinates$Population %in% map_aripo ~ "Aripo",
  Coordinates$Population %in% map_quare ~ "Quare",
  Coordinates$Population %in% map_tacarigua ~ "Tacarigua",
  Coordinates$Population %in% map_tunapuna ~ "Tunapuna",
  Coordinates$Population %in% map_arouca ~ "Arouca",
  Coordinates$Population %in% map_guanapo ~ "Guanapo",
  Coordinates$Population %in% map_arima ~ "Arima",
  Coordinates$Population %in% map_elcedro ~ "El Cedro",
  Coordinates$Population %in% map_turure ~ "Turure",
  Coordinates$Population %in% map_oropuche ~ "Oropuche",
  Coordinates$Population %in% map_yarra ~ "Yarra",
  Coordinates$Population %in% map_marianne ~ "Marianne",
  Coordinates$Population %in% map_paria ~ "Paria",
  Coordinates$Population %in% map_damier ~ "Damier",
  Coordinates$Population %in% map_other ~ "Other"
)
# this codes the poptype_broad column to indicate our system knowledge that certain watersheds were not colonized by guppies until researchers' experiments
Coordinates$Poptype_broad<-ifelse(Coordinates$Population %in% c("Damier", "Turure"), "Introduction",
                               ifelse((Coordinates$Population %in% c("El Cedro") & Coordinates$Predation=="Low"), "Introduction", 
                                      ifelse((Coordinates$Population %in% c("El Cedro") & Coordinates$Predation=="Low"), "Introduction", 
                                             ifelse((Coordinates$Population %in% c("Arima") & Coordinates$Predation=="High" & Coordinates$Collection_start>2001), "Introduction", Coordinates$Poptype))))

#write_csv(Coordinates, path = "~/Documents/Projects/R scripts/Meta Analysis/Data/Map/site_type.csv")

```

### Follow Up on reviewer comment

```{r anova resample loop}

dat<-FALSE 

i <- 1

anova_loop.resample <- function(dat=dat){
  
  trait <- as.factor(dat$TraitID)

#output.perm <- array(dim=c(1000, 3, (length(unique(dat$TraitID)))))
output.perm <- array(dim=c(100, 3, (length(unique(dat$TraitID)))))

  
  for (i in 1: length(levels(trait))){
    dat.temp <- dat[dat$TraitID==levels(trait)[i],]
    print(i)
    traitID <- levels(trait)[i]
    
    #resampling permutations

output.permj <- matrix (ncol=3, nrow=(100))
colnames(output.permj) <- c("TraitID", "R2.RE", "RadjRE")

for (j in 1:100){

  dat.temp$PredationRESAMPLE <- sample(dat.temp$Predation, replace = FALSE)
  
  aov.modRE   <- aov(MeanValue~PredationRESAMPLE,data=dat.temp)
  anova.modRE <- stats:::anova.lm(aov.modRE)
  mod.ssRE    <- anova.modRE$"Sum Sq"
  mod.pesRE   <- mod.ssRE/(mod.ssRE+mod.ssRE[length(mod.ssRE)])          # calculate pes
  
  rRE <- summary.lm(aov.modRE)                                     # summary of linear model
  
  R2.RE     <- round(rRE$"r.squared",3)                              # R squared
  R_adjRE <- round(rRE$"adj.r.squared",3)
  

  #output.RE[j,] <- c(R,R_adj,partial_eta,F_value,p_value, traitID, justr, R2.RE, R_adjRE) 
  output.permj[j,] <- c(traitID, R2.RE, R_adjRE)
}
 output.perm[,,i] <- output.permj   

  }
output.perm<-as.data.frame(output.perm)
colnames(output.perm)<- rep(c("TraitID", "R2", "R2adj"),len = (ncol(output.perm))) #rename the columns as data type

#reformats the output as a nice df
output.perm<-pivot_longer(output.perm, cols = 1:3, names_to = 'type') 
output.perm<-pivot_wider(output.perm, names_from = type, values_from = value)

resampleR2<-data.frame(TraitID = (unlist(output.perm$TraitID)), R2 = (unlist(output.perm$R2)), R2adj = (unlist(output.perm$R2adj))) %>%
  mutate(R2 = as.numeric(R2), TraitID = as.numeric(TraitID), R2adj = as.numeric(R2adj))
return(resampleR2)

}

```




```{r loading all dfs}
# load in all data tables as unique outputs
#The trait filter groups are in Tallies.rmd, commented the corresponding df
output_all<-read.csv(paste(wd,'/Data/TraitR2_among.csv',sep="")) #new.data
output_south<-read.csv(paste(wd,'/Data/TraitR2_South.csv',sep="")) #traits_s
output_caroni<-read.csv(paste(wd,'/Data/TraitR2_Caroni.csv',sep=""))#traits_d
output_south_drainages<-read.csv(paste(wd,'/Data/TraitR2_Among_Drainage.csv',sep="")) #traits_d_all
output_intros<-read.csv(paste(wd,'/Data/TraitR2_intro.csv',sep="")) #traits_i
output_intros_broad<-read.csv(paste(wd,'/Data/TraitR2_intro_broad.csv',sep="")) #traits_i_broad

resample_all<-anova_loop.resample(dat = new.data)
resample_south<-anova_loop.resample(dat = traits_s)
resample_caroni<-anova_loop.resample(dat = traits_d)
resample_south_drainages<-anova_loop.resample(dat = traits_d_all)
resample_intros<-anova_loop.resample(dat = traits_i)
resample_intros_broad<-anova_loop.resample(dat = traits_i_broad)
```


```{r function to create one output}

full_output<-function(dat = dat, resampleDat = resampleDat){
    
  #calculate the mean R2 per trait
R2means<-resampleDat %>% 
  group_by(TraitID) %>% 
  summarize(R2mean = mean(R2))

  #create output df with all R2 data
rDat <- left_join(dat, R2means, by = 'TraitID') #merge by TraitID (same col name)
    
  #run model with original and mean perm R2
ex.mod <- lm(R.2 ~ R2mean, data=rDat)
resid <- residuals(ex.mod)
rDat$resid <- resid
  
return(rDat)
}

Rdat_all<-full_output(dat = output_all, resampleDat = resample_all)
Rdat_south<-full_output(dat = output_south, resampleDat = resample_south)
Rdat_caroni<-full_output(dat = output_caroni, resampleDat = resample_caroni)
Rdat_south_drainages<-full_output(dat = output_south_drainages, resampleDat = resample_south_drainages)
Rdat_intros<-full_output(dat = output_intros, resampleDat = resample_intros)
#Rdat_intros_broad<-full_output(dat = output_intros_broad, resampleDat = resample_intros_broad)


#issues with intros_broad (dropping a data point in the model? 218 instead of 219) I will troubleshoot
  #calculate the mean R2 per trait
means_i<-resample_intros_broad %>% 
  group_by(TraitID) %>% 
  summarize(R2mean = mean(R2))

  #create output df with all R2 data
intros_rdat <- left_join(output_intros_broad, means_i, by = 'TraitID') #merge by TraitID (same col name)
    
  #run model with original and mean perm R2
ex.mod <- lm(R.2 ~ R2mean, data=intros_rdat)
resid <- residuals(ex.mod)
intros_rdat$resid <- resid #replacement has 218 rows, data has 219????

```

```{r visualization}
# for all the traits
plot(Rdat_all$R2mean, Rdat_all$R.2)
hist(Rdat_all$R2mean)
hist(Rdat_all$resid)

plot_a <- ggplot(Rdat_all, aes(R2mean, R.2))

plot_a + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R²") + xlab("Mean R² expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

# for south

plot(Rdat_south$R2mean, Rdat_south$R.2)
hist(Rdat_south$R2mean)
hist(Rdat_south$resid)

plot_s <- ggplot(Rdat_south, aes(R2mean, R.2))

plot_s + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R²") + xlab("Mean R² expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

# for south_drainages

plot(Rdat_south_drainages$R2mean, Rdat_south_drainages$R.2)
hist(Rdat_south_drainages$R2mean)
hist(Rdat_south_drainages$resid)

plot_sd <- ggplot(Rdat_south_drainages, aes(R2mean, R.2))

plot_sd + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R²") + xlab("Mean R² expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

# for caroni
plot(Rdat_caroni$R2mean, Rdat_caroni$R.2)
hist(Rdat_caroni$R2mean)
hist(Rdat_caroni$resid)

plot_c <- ggplot(Rdat_caroni, aes(R2mean, R.2))

plot_c + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R²") + xlab("Mean R² expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

# for all the traits
plot(Rdat_intros$R2mean, Rdat_intros$R.2)
hist(Rdat_intros$R2mean)
hist(Rdat_intros$resid)

plot_i <- ggplot(Rdat_intros, aes(R2mean, R.2))

plot_i + geom_point(size=1) + #geom_smooth(method="lm") + 
  geom_abline(intercept = 0, slope = 1) +
  ylab("Actual R²") + xlab("Mean R² expected at random") +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  #expand_limits(x=c(0,1)) + 
  #mytheme_bw() + 
  #scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  theme(plot.margin=unit(c(0.3,0.4,0.3,0.3),"cm"))

```





