---
title: "Tallies"
output: html_notebook
---

```{r read in}
library(tidyverse)
library(ggplot2)
#setwd("<YOURPATHWAY>")
wd <- getwd()
new.data<- read.csv(paste(wd,'/Data/MetaData.csv',sep=""), header=TRUE, sep=",") #MetaData saved feb 11th
data<- read.csv(paste(wd,'/Data/Data.csv',sep=""), header=TRUE, sep=",")%>% #Data.csv is saved .xlsx feb 10th
  select(StudyID:Comments) 
r.squared<- read.csv(paste(wd,'/Data/TraitR2_among.csv',sep=""), header=TRUE, sep=",") #R2 generated feb 11th 
```

Tallies and Counts
## Effect Size
using only traits we can calculate cohen's d for...
```{r numbers for cohens d}
#traits with data needed for cohen's d
test<-new.data %>% 
  filter(!is.na(MeanValue)) %>% 
  filter(!is.na(Number)) %>% 
  filter(!is.na(SD) | !is.na(SE))

# drainage level summary information, idk why this is here
new.data[!duplicated(new.data$TraitID),] %>% 
  group_by(Slope.type) %>%
  summarize(count = n())

test<-new.data %>% 
  filter(Slope.type == "South") %>% 
  group_by(TraitID) %>% 
  mutate(unique_drainages = n_distinct(Drainage))

test$Drainage.type<-ifelse(test$unique_drainages==1, test$Drainage, "Mixed")

test[!duplicated(test$TraitID),] %>% 
  group_by(Drainage.type) %>%
  summarize(count = n())
  
```


```{r calculate the d}
library(effsize)

test<-new.data %>%
  filter(!is.na(MeanValue)) %>% 
  filter(!is.na(Number)) %>% 
  filter(!is.na(SD) | !is.na(SE)) %>%  
  group_by(TraitID) %>% 
  filter(n_distinct(Poptype)>1) %>% 
  select(1:3,20,14,19,21:23,38,4:5,10,25:31,37,40,41)

good_trait<-test %>% 
  filter(Poptype=="Introduction" & !is.na(Source_pop) & !Source_pop =="Unknown") %>% 
  distinct(TraitID) %>% 
  pull(TraitID) #turns to vector, 114 traits with explicit source pops listed

test<-test %>% 
  filter(TraitID %in% good_trait)

source_pops<-pull(test, Source_pop)

test<-test %>% 
  filter(PopulationRaw %in% source_pops | !is.na(Source_pop)) %>% 
  mutate(type = ifelse(is.na(Source_pop), "Source", "Intro")) %>% 
  mutate(pop.temp = ifelse(!is.na(Source_pop), Source_pop, PopulationRaw)) %>% 
  group_by(TraitID, pop.temp) %>% 
  mutate(introID = cur_group_id()) %>% 
  arrange(introID) %>% 
  filter(n_distinct(type)>1)

test$MeanValue<-as.numeric(test$MeanValue)
test$pop.temp<-as.factor(test$pop.temp)
test$type<-as.factor(test$type)
test$introID<-as.factor(test$introID)
test2<-test %>% 
  ungroup() %>% 
  slice_tail(n = 20) %>% 
  filter(!introID %in% c(97,82))

test2$introID<-as.factor(test2$introID) %>% 
  droplevels()

mean<-pull(test2, MeanValue)
type<-pull(test2, type)
id<-pull(test2, introID)
cohen.d(d = mean, f = type, subject = id, paired = TRUE, data = test2) #this isn't really working and also needs real data not meta data

library(esc) # this uses metaData but unfortunately we don't have a large enough metadata sample to do the pop vs intro questions this way (n= 3)

test3<-test %>% 
  select(4,18:21,24:26) %>% 
  ungroup() %>% 
  slice_tail(n = 20) %>% 
  filter(!introID == 97)

test3<-pivot_wider(test3, names_from = type, values_from = c(MeanValue, SD, SE, Number))

esc_mean
```


## North vs South Slope
Looked at the ratio of north to south in traits with both populations (histogram)
Also looked at how many traits have over 4 populations on both the north and south slope (48)
```{r ratio of north and south slope pops}
slope.ratio<-new.data %>% 
  group_by(TraitID) %>% 
  summarise(South = sum(Slope=="South"), North = sum(Slope=="North"), ratio = North/South)
#visualize only mixed slopes traits by the ratio of north to south, are they evenly mixed or mostly one?
slope.ratio %>% 
  subset(ratio > 0 & !(ratio == Inf)) %>% 
  ggplot(aes(x = ratio)) + geom_histogram()

#can we use a subset to look at traits that have enough north and south to have a R2 for each?
test<-slope.ratio %>% 
  filter((South>4 & North>4) |  (North == 0 | South == 0 )) #302 with over 4 in each slope or only one slope

 #231 south only with 4+ pops
 #29 north only with 4+ pops
traits<-test %>% 
  filter(South>4 & North>4) %>% 
  select(TraitID)

traits<-left_join(traits, new.data, by = "TraitID")
summary(as.factor(traits$StudyID)) #only 2 studies :(
```


```{r caroni and oropuche question }

#Caroni and Oropuche Question, how many traits have more than 2 hp and lp in both oropuche and caroni?
traits_s<-new.data %>% 
  group_by(TraitID, Drainage, Predation) %>% 
  tally() %>% 
  filter(n>2) %>% 
  filter(Drainage == "Caroni" | Drainage == "Oropuche")

traits_s %>% 
  group_by(TraitID) %>% 
  tally() %>% 
  filter(n>3) #so there's 10 traits with 2 hp and lp in both caroni and oropuche, not appropriate for anova
```

## Introductions

We need to tag sites that are known introductions (Damier, Turure, El Cedro, Arima post 2001 etc) with a stage different from studies that knowingly compare intro sites to their source population
```{r Introduction counts}
#just to start, how many traits have an introduction
new.data %>% 
  group_by(StudyID, TraitID) %>% 
  filter(Poptype=="Introduction") %>% 
  count(StudyID)  %>%
  tally() 
  
#there's 111 traits with at least one introduction in them, 11 StudyIDs
```

## Subsets for Among and Within

rewrite traits_s or traits_d as new.data and run through anova loop!
```{r subsets for among and within questions}
# How many pops have more than 2 hp and lp in south
ids<-new.data %>%
  group_by(TraitID, Slope, Predation) %>%
  tally() %>% #tallies how many populations in high and low for each Slope within a trait
  group_by(TraitID) %>% 
  filter(length(unique(Slope))>1) %>% #selects traits with multiple slopes
  filter(Slope == "South") %>% #select only South pops to determine which traits are viable
  filter(n>1) %>%
  filter(length(unique(Predation))>1) %>% #selects for traits with natural HP and LP with 2+ pops
  filter(!duplicated(TraitID)) %>% 
  select(TraitID)

traits_s<-new.data %>% 
  filter(TraitID %in% ids$TraitID) %>% 
  filter(Slope== "South") # 204 traits, done and on drive

# How many pops have more than 2 hp and lp in caroni, and their own comparison R2 prep (excluding north pops from those traits if there are any)
ids_south<-new.data %>%
  group_by(TraitID, Slope, Predation) %>%
  tally() %>% 
  group_by(TraitID) %>% 
  filter(Slope == "South") %>%
  filter(n>1) %>%
  filter(length(unique(Predation))>1) %>%
  filter(!duplicated(TraitID)) %>% 
  select(TraitID) #432 traits total, use this to select traits with enough south pops to be viable for drainage analysis

ids_d<-new.data %>%
  filter(TraitID %in% ids_south$TraitID) %>% 
  filter(Slope == "South") %>% #only use south pops
  group_by(TraitID, Drainage, Predation) %>%
  tally() %>% #tallies how many populations in high and low for each driange within a trait
  group_by(TraitID) %>% 
  filter(length(unique(Drainage))>1) %>% #selects traits with both drainages
  filter(Drainage == "Caroni") %>% #select only caroni pops to determine which traits are viable
  filter(n>1) %>%
  filter(length(unique(Predation))>1) %>% #selects for traits with natural HP and LP with 2+ pops
  filter(!duplicated(TraitID)) %>% 
  select(TraitID)

traits_d<-new.data %>%
  filter(TraitID %in% ids_d$TraitID) %>% #use only traits that have enough caroni pops
  filter(Drainage == "Caroni") # 317 traits

#this will be used as the among drainage comparisons to control for variation between slopes
traits_d_all<-new.data %>%
  filter(TraitID %in% ids_d$TraitID) %>% #use only traits that have enough caroni pops
  filter(Slope == "South")

#Introductions
ids_i<-new.data %>% 
  group_by(TraitID, Poptype, Predation) %>%
  tally() %>% #tallies how many populations in high and low for intro or natural pops per trait
  group_by(TraitID) %>% 
  filter(length(unique(Poptype))>1) %>% #selects traits with both natural and introduction populations
  filter(Poptype == "Natural") %>% #select only natural pops to determine which traits are viable
  filter(n>1) %>%
  filter(length(unique(Predation))>1) %>% #selects for traits with natural HP and LP with 2+ pops
  filter(!duplicated(TraitID)) %>% 
  select(TraitID)

traits_i<-new.data %>% 
  filter(TraitID %in% ids_i$TraitID) %>% 
  filter(Poptype == "Natural") #232 traits, done and on drive
```
##

Visualization and Plots

## R2 plots (old school visualization)
```{r R2 prep}
new.data<-left_join(new.data,r.squared, by = "TraitID")
new.data$R.2<-as.numeric(new.data$R.2)
```



```{r plots violin}
ggplot(data = new.data, aes(x=TraitType2, y = R.2, fill = TraitType2)) + geom_violin() 
ggplot(data = new.data, aes(x=TraitType1, y = R.2, fill = TraitType1)) + geom_violin() 

new.data %>% 
  filter(unique_slopes==1) %>% 
ggplot(aes(x=Slope, y = R.2, fill = Slope)) + geom_violin() 

new.data %>% 
  filter(unique_slopes==1) %>% 
ggplot(aes(x=TraitType2, y = R.2, fill = Slope)) + geom_violin()

new.data %>% 
  filter(unique_slopes==1) %>% 
ggplot(aes(x=TraitType1, y = R.2, fill = Slope)) + geom_violin()

#ggplot(data = new.data, aes(x=StudyType, y = R.2, fill = StudyType)) + geom_violin() 

#sex specific plots
new.data %>% 
  filter(Sex == "M" | Sex == "F") %>% 
ggplot( aes(x = Sex, y = R.2, fill = Sex)) + geom_violin()

new.data %>% 
  filter(Sex == "M" | Sex == "F") %>% 
ggplot( aes(x = TraitType2, y = R.2, fill = Sex)) + geom_violin()

new.data %>% 
  filter(Sex == "M" | Sex == "F") %>% 
ggplot( aes(x = TraitType1, y = R.2, fill = Sex)) + geom_violin()

```
##


