---
title: "Tallies"
output: html_notebook
---

```{r read in}
library(tidyverse)
library(ggplot2)
#setwd("<YOURPATHWAY>")
wd <- getwd()
new.data<- read.csv(paste(wd,'/Data/MetaData.csv',sep=""), header=TRUE, sep=",") #MetaData saved feb 11th
data<- read.csv(paste(wd,'/Data/Data.csv',sep=""), header=TRUE, sep=",")%>% #Data.csv is saved .xlsx feb 10th
  select(StudyID:Comments) 
r.squared<- read.csv(paste(wd,'/Data/TraitR2.csv',sep=""), header=TRUE, sep=",") #R2 generated feb 11th 
```

Tallies and Counts
## Effect Size
using only traits we can calculate cohen's d for...
```{r numbers for cohens d}
#traits with data needed for cohen's d
test<-new.data %>% 
  filter(!is.na(MeanValue)) %>% 
  filter(!is.na(Number)) %>% 
  filter(!is.na(SD) | !is.na(SE))

new.data[!duplicated(new.data$TraitID),] %>% 
  group_by(Slope.type) %>%
  summarize(count = n())

test<-new.data %>% 
  filter(Slope.type == "South") %>% 
  group_by(TraitID) %>% 
  mutate(unique_drainages = n_distinct(Drainage))

test$Drainage.type<-ifelse(test$unique_drainages==1, test$Drainage, "Mixed")

test[!duplicated(test$TraitID),] %>% 
  group_by(Drainage.type) %>%
  summarize(count = n())
  
```

## North vs South Slope
Looked at the ratio of north to south in traits with both populations (histogram)
Also looked at how many traits have over 4 populations on both the north and south slope (48)
```{r ratio of north and south slope pops}
slope.ratio<-new.data %>% 
  group_by(TraitID) %>% 
  summarise(South = sum(Slope=="South"), North = sum(Slope=="North"), ratio = North/South)
#visualize only mixed slopes traits by the ratio of north to south, are they evenly mixed or mostly one?
slope.ratio %>% 
  subset(ratio > 0 & !(ratio == Inf)) %>% 
  ggplot(aes(x = ratio)) + geom_histogram()

#can we use a subset to look at traits that have enough north and south to have a R2 for each?
test<-slope.ratio %>% 
  filter((South>4 & North>4) |  (North == 0 | South == 0 )) #302 with over 4 in each slope or only one slope

 #231 south only with 4+ pops
 #29 north only with 4+ pops
traits<-test %>% 
  filter(South>4 & North>4) %>% 
  select(TraitID)

traits<-left_join(traits, new.data, by = "TraitID")
summary(as.factor(traits$StudyID)) #only 2 studies :(
```


```{r caroni and oropuche question }

#Caroni and Oropuche Question, how many traits have more than 2 hp and lp in both oropuche and caroni?
traits_s<-new.data %>% 
  group_by(TraitID, Drainage, Predation) %>% 
  tally() %>% 
  filter(n>2) %>% 
  filter(Drainage == "Caroni" | Drainage == "Oropuche")

traits_s %>% 
  group_by(TraitID) %>% 
  tally() %>% 
  filter(n>3) #so there's 10 traits with 2 hp and lp in both caroni and oropuche, not appropriate for anova
```

## Introductions

We need to tag sites that are known introductions (Damier, Turure, El Cedro, Arima post 2001 etc) with a stage different from studies that knowingly compare intro sites to their source population
```{r Introduction counts}
#just to start, how many traits have an introduction
new.data %>% 
  group_by(StudyID, TraitID) %>% 
  filter(Poptype=="Introduction") %>% 
  count(StudyID)  %>%
  tally() 
  
#there's 111 traits with at least one introduction in them, 11 StudyIDs
```

## Subsets for Among and Within

rewrite traits_s or traits_d as new.data and run through anova loop!
```{r subsets for among and within questions}
# How many pops have more than 2 hp and lp in south
ids<-new.data %>%
  group_by(TraitID, Slope, Predation) %>%
  tally() %>% #tallies how many populations in high and low for each Slope within a trait
  group_by(TraitID) %>% 
  filter(length(unique(Slope))>1) %>% #selects traits with multiple slopes
  filter(Slope == "South") %>% #select only South pops to determine which traits are viable
  filter(n>1) %>%
  filter(length(unique(Predation))>1) %>% #selects for traits with natural HP and LP with 2+ pops
  filter(!duplicated(TraitID)) %>% 
  dplyr::select(TraitID)

traits_s<-new.data %>% 
  filter(TraitID %in% ids$TraitID) %>% 
  filter(Slope== "South") # 204 traits, done and on drive

# How many pops have more than 2 hp and lp in caroni, and their own comparison R2 prep (excluding north pops from those traits if there are any)
ids_south<-new.data %>%
  group_by(TraitID, Slope, Predation) %>%
  tally() %>% 
  group_by(TraitID) %>% 
  filter(Slope == "South") %>%
  filter(n>1) %>%
  filter(length(unique(Predation))>1) %>%
  filter(!duplicated(TraitID)) %>% 
  dplyr::select(TraitID) #432 traits total, use this to select traits with enough south pops to be viable for drainage analysis

ids_d<-new.data %>%
  filter(TraitID %in% ids_south$TraitID) %>% 
  filter(Slope == "South") %>% #only use south pops
  group_by(TraitID, Drainage, Predation) %>%
  tally() %>% #tallies how many populations in high and low for each driange within a trait
  group_by(TraitID) %>% 
  filter(length(unique(Drainage))>1) %>% #selects traits with both drainages
  filter(Drainage == "Caroni") %>% #select only caroni pops to determine which traits are viable
  filter(n>1) %>%
  filter(length(unique(Predation))>1) %>% #selects for traits with natural HP and LP with 2+ pops
  filter(!duplicated(TraitID)) %>% 
  dplyr::select(TraitID)

traits_d<-new.data %>%
  filter(TraitID %in% ids_d$TraitID) %>% #use only traits that have enough caroni pops
  filter(Drainage == "Caroni") # 317 traits

#this will be used as the among drainage comparisons to control for variation between slopes
traits_d_all<-new.data %>%
  filter(TraitID %in% ids_d$TraitID) %>% #use only traits that have enough caroni pops
  filter(Slope == "South")

#Introductions
ids_i<-new.data %>% 
  group_by(TraitID, Poptype, Predation) %>%
  tally() %>% #tallies how many populations in high and low for intro or natural pops per trait
  group_by(TraitID) %>% 
  filter(length(unique(Poptype))>1) %>% #selects traits with both natural and introduction populations
  filter(Poptype == "Natural") %>% #select only natural pops to determine which traits are viable
  filter(n>1) %>%
  filter(length(unique(Predation))>1) %>% #selects for traits with natural HP and LP with 2+ pops
  filter(!duplicated(TraitID)) %>% 
  dplyr::select(TraitID)

traits_i<-new.data %>% 
  filter(TraitID %in% ids_i$TraitID) %>% 
  filter(Poptype == "Natural") #209 traits, done and on drive

ids_i_broad<-new.data %>% 
  group_by(TraitID, Poptype_broad, Predation) %>%
  tally() %>% #tallies how many populations in high and low for intro or natural pops per trait
  group_by(TraitID) %>% 
  filter(length(unique(Poptype_broad))>1) %>% #selects traits with both natural and introduction populations
  filter(Poptype_broad == "Natural") %>% #select only natural pops to determine which traits are viable
  filter(n>1) %>%
  filter(length(unique(Predation))>1) %>% #selects for traits with natural HP and LP with 2+ pops
  filter(!duplicated(TraitID)) %>% 
  dplyr::select(TraitID)

traits_i_broad<-new.data %>% 
  filter(TraitID %in% ids_i_broad$TraitID) %>% 
  filter(Poptype_broad == "Natural") #219 traits, done and on drive
```



##


```{r temp}
library(tidyverse)
dat <- read_excel("~/Downloads/metaanalysis_finaltraits.xlsx")
dat<-dat %>% 
  arrange(StudyID, StudyType, Trait, Sex, Predation)
sex.length<-rle(as.character(paste(dat$Trait,dat$Sex,dat$StudyType)))  
dat$TraitID<-rep(seq_along(sex.length$lengths), sex.length$lengths)

test<-dat %>% 
  group_by(Kingsolver_traits) %>% 
  subset(!duplicated(TraitID))

summary(as.factor(test$Kingsolver_traits))

x<-subset(test, test$Kingsolver_traits=="Behaviour")


```

